Comb associative operads

### 2018-09-03 CAs_operads.tex
MAJ partie 3.3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{The \texorpdfstring{$3$}{3}-comb associative operad}
\label{subsubsec:CAs_3}
Following the Buchberger algorithm, we need to put the right
members of~\eqref{equ:branching_pair_CAs_3} in relation by $\Rew$. To
respect the lexicographic property of the prefix words, this leads to
the new relation
\begin{equation} \label{equ:rew_2}
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-3.60){};
        \node(2)at(2.00,-5.40){};
        \node(4)at(4.00,-7.20){};
        \node(6)at(6.00,-7.20){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(7);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.5){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \enspace \Rew \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.20,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-4.50){};
        \node(4)at(4.00,-4.50){};
        \node(6)at(6.00,-6.75){};
        \node(8)at(8.00,-6.75){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-4.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(3.00,1.74){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}\,.
\end{equation}
The Buchberger algorithm applied on binary trees of degrees $5$, $6$,
and $7$ provides the new relations \smallbreak
\input{Buchberger_relations_5-7}
\medbreak

\begin{Theorem} \label{thm:convergent_rewrite_rule_CAs_3}
    The set $\Rew$ of rewrite rules containing~\eqref{equ:rew_1},
    \eqref{equ:rew_2}, \eqref{equ:rew_3}---\eqref{equ:rew_11} is a
    convergent presentation of~$\CAs{3}$.
\end{Theorem}
\begin{proof}
    Let us show that the rewrite relation $\RewContext$ induced by
    $\Rew$ is convergent. First, for every relation $\Tfr \Rew \Tfr'$,
    we have $\Tfr > \Tfr'$. Therefore, by
    Lemma~\ref{lem:prefix_word_termination}, $\RewContext$ is
    terminating. Moreover, the greatest degree of a tree appearing in
    $\Rew$ is~$7$ so that, from Lemma~\ref{lem:degree_confluence}, to
    show that $\RewContext$ is convergent, it is enough to prove that
    each tree of degree at most $13$ admits exactly one normal form.
    Equivalently, this amounts to show that the number of normal forms
    of trees of arity $n$ is equal to $\#\CAs{3}(n)$. By computer
    exploration, we get the same sequence
    \begin{equation} \label{equ:dimensions_CAs_3}
        1, 1, 2, 4, 8, 14, 20, 19, 16, 14, 14, 15, 16, 17
    \end{equation}
    for $\#\CAs{3}(n)$ and for the numbers of normal forms of arity $n$,
    when $ 1 \leq n \leq 14$. Hence, we get our following main result.
\end{proof}
\medbreak

@@

\begin{center}
  \begin{math}
    \begin{array}{|c|l|}
      \hline
      \gamma & \mbox{The sequence of first cardinalities of } 
       \CAs{\gamma} \mbox{completions}\\ \hline
      3 & 0, 0, 0, 1, 1, 2, 3, 4, 0, 0, \dots \\ \hline
      4 & 0, 0, 0, 0, 1, 1, 0, 3, 4, 5, 18, 22, 11, 12, 15, 19, 25, 36,
       44, 52, 68, 79, 93, 105, 106, 109 \\ \hline
      5 & 0, 0, 0, 0, 0, 1, 1, 0, 0, 4, 5, 8, 18, 31, 36, 48, 73, 111,
       172, 272, 455, 783 \\ \hline
      6 & 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 5, 6, 11, 23, 30, 48, 73, 
      117, 204, 348, 589, 1004 \\ \hline
      7 & 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 6, 7, 16, 24, 32, 49, 
      88, 150, 261, 475, 854\\ \hline
      8 & 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 7, 8, 21, 29, 
      34, 53, 93, 172, 311, 565\\ \hline
      9 & 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 0, 8, 9, 28, 
      30, 36, 57, 101, 185, 348, 648\\ \hline
    \end{array}
  \end{math}
\end{center}

From these computer explorations, we conjecture that this algorithm
does not provide a finite convergent presentation of $\CAs{\gamma}$, 
when $\gamma \geq 4$.

@@

\Todo{résultats}

We conjecture that there is no finite and convergent presentation
of $\CAs{\gamma}$ with one generator when $\gamma \geq 4$.
We discuss the cases with more then one generator in the perspectives.

%By computer exploration, we have already some information in this
%direction: we have the sequence
%\begin{equation}
    %1, 1, 2, 5, 13, 35, 96, 264, 724, 1973, 5335, 14390, 38872, 105141,
    %284929, 774254
%\end{equation}
%for the first dimensions for $\CAs{4}$. Moreover, for $\CAs{5}$, 
%we get the
%sequence
%\begin{equation}
%1, 1, 2, 5, 14, 41, 124, 384, 1210, 3861, 12440, 40392, 131997,
%433782, 1432696, 4752857
%\end{equation}
%of dimensions and the first cardinalities
%\begin{math}
    %0, 0, 0, 0, 0, 1, 1, 0, 0, 4, 5
%\end{math}
%for any convergent orientation of $\CongrCAs{5}$. Finally, for
%$\CAs{6}$, we get the sequence
%\begin{equation}
   %1, 1, 2, 5, 14, 42, 131, 420, 1375, 4576, 15431, 52598,
   %180895, 626862, 2186504, 7670138
%\end{equation}
%of dimensions and the first cardinalities
%\begin{math}
    %0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0
%\end{math}
%for any convergent orientation of $\CongrCAs{6}$. We can notice that
%only $\CAs{3}$ seems to have oscillating first dimensions.
%\medbreak

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{The \texorpdfstring{$3$}{3}-comb associative operad}
\label{subsubsec:CAs_3}

The Buchberger algorithm applied on binary trees of degrees $4$ to $7$ 
provides the new relations \smallbreak
\input{Buchberger_relations_CAs3}
\medbreak

\begin{Theorem} \label{thm:convergent_rewrite_rule_CAs_3}
    The set $\Rew$ of rewrite rules containing~\eqref{equ:rew_1},
    and \eqref{equ:rew_2}---\eqref{equ:rew_11} is a
    convergent presentation of~$\CAs{3}$.
\end{Theorem}
\begin{proof}
    Let us show that the rewrite relation $\RewContext$ induced by
    $\Rew$ is convergent. First, for every relation $\Tfr \Rew \Tfr'$,
    we have $\Tfr > \Tfr'$. Therefore, by
    Lemma~\ref{lem:prefix_word_termination}, $\RewContext$ is
    terminating. Moreover, the greatest degree of a tree appearing in
    $\Rew$ is~$7$ so that, from Lemma~\ref{lem:degree_confluence}, to
    show that $\RewContext$ is convergent, it is enough to prove that
    each tree of degree at most $13$ admits exactly one normal form.
    Equivalently, this amounts to show that the number of normal forms
    of trees of arity $n$ is equal to $\#\CAs{3}(n)$. By computer
    exploration, we get the same sequence
    \begin{equation} \label{equ:dimensions_CAs_3}
        1, 1, 2, 4, 8, 14, 20, 19, 16, 14, 14, 15, 16, 17
    \end{equation}
    for $\#\CAs{3}(n)$ and for the numbers of normal forms of arity $n$,
    when $ 1 \leq n \leq 14$. Hence, we get our following main result.
\end{proof}
\medbreak

@@

\begin{center}
  \begin{math} \label{arr:completions_CAs}
    \begin{array}{|c|l|} 
      \hline
      \gamma & \mbox{The sequence of first cardinalities of } 
       \CAs{\gamma} \mbox{completions}\\ \hline
      3 & 0, 0, 0, 1, 1, 2, 3, 4, 0, 0, \dots \\ \hline
      4 & 0, 0, 0, 0, 1, 1, 0, 3, 4, 5, 18, 22, 11, 12, 15, 19, 25, 36,
       44, 52, 68, 79, 93, 105, 106, 109 \\ \hline
      5 & 0, 0, 0, 0, 0, 1, 1, 0, 0, 4, 5, 8, 18, 31, 36, 48, 73, 111,
       172, 272, 455, 783 \\ \hline
      6 & 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 5, 6, 11, 23, 30, 48, 73, 
      117, 204, 348, 589, 1004 \\ \hline
      7 & 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 6, 7, 16, 24, 32, 49, 
      88, 150, 261, 475, 854\\ \hline
      8 & 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 7, 8, 21, 29, 
      34, 53, 93, 172, 311, 565\\ \hline
      9 & 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 0, 8, 9, 28, 
      30, 36, 57, 101, 185, 348, 648\\ \hline
    \end{array}
  \end{math}
\end{center}
\medbreak

From these computer explorations, we conjecture that this algorithm
does not provide a finite convergent presentation of $\CAs{\gamma}$, 
when $\gamma \ge 4$.

@@

We did not found a finite presentation for $\CAs{4}$, $\CAs{5}$ and 
$\CAs{6}$ until arity $12$. 
We conjecture that there is no finite and convergent presentation
of $\CAs{\gamma}$ with one generator when $\gamma \geq 4$.
We discuss the cases with more then one generator in the perspectives.
\medbreak

Thanks to the partial completions we compute in table
\eqref{arr:completions_CAs}, we can compute the following first 
dimensions of $\CAs{\gamma}$.
\begin{center}
  \begin{math} \label{arr:dimensions_CAs}
    \begin{array}{|c|l|} 
      \hline
      \gamma & \mbox{The sequence of first cardinalities of } 
       \CAs{\gamma}\\ \hline
      4 & 1, 1, 2, 5, 13, 35, 96, 264, 724, 1973, 5335, 14390, 38872,
          105141, 284929, 774254, 2111088,\\
        & 5778420, 15882100, 43837388, 121507230 \\ \hline
      5 & 1, 1, 2, 5, 14, 41, 124, 384, 1210, 3861, 12440, 40392, 131997,
          433782, 1432696, 4752857, \\ 
        & 15829261, 52905635, 177395246 \\ \hline
      6 & 1, 1, 2, 5, 14, 42, 131, 420, 1375, 4576, 15431, 52598, 180895,
          626862, 2186504, 7670138, \\ 
        & 27041833, 95764569 \\ \hline
      7 & 1, 1, 2, 5, 14, 42, 132, 428, 1420, 4796, 16432, 56966, 199444,
          704140, 2503914, 8959699, \\
        & 32236657, 116551168\\ \hline
      8 & 1, 1, 2, 5, 14, 42, 132, 429, 1429, 4851, 16718, 58331, 205632,
          731272, 2620176, 9449688, \\
        & 34276116, 124958386\\ \hline
      9 & 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4861, 16784, 58695, 207452,
          739840, 2658936, 9620232, \\
        & 35011566, 128082515\\ \hline
    \end{array}
  \end{math}
\end{center}
